Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-03-05 WRITEUP/ms.tex
edit def of T; updated abstract

\begin{abstract}
We study the dynamical properties of molecular clouds complexes (MC) in a prototypical galaxy (\flower)
and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
We identify MCs using H$_2$ density, 
and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$)
with those observed in nearby and \z$\sim$\,2 galaxies.
The MCs span a mass range of $10^{6.5-9}$\,\Msun. Their velocity dispersions and gas surface densities are 
systematically higher than those observed in nearby galaxies but comparable to 
starburst galaxies, with $\sigma\gtrsim$\,100\,\kms.
This likely results from the strong supernova and stellar feedback \flower experienced over multiple episodes of
bursty \SF and cosmic gas inflows.
Our results suggest that the more massive and bigger MCs of \flower compared to the Milky Way
manifest from the higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
Virial analysis indicates that the MCs in the main disk of \flower are unbound, but its substructures have lower virial parameters,
consistent with the notion that collapsing structures result from gravitational instability within globally stable structures.
% , which are supported by turbulence and rotation on large scale.
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result from the weaker stellar feedback
due to the less frequent \SF compared to \flower.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the
Next Generation Very Large Array (ngVLA)
will furnish useful data to test our findings and simulation to shed
light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

In theoretical and observational works, gravitational instability has been widely discussed in the 
context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}, or the Toomre Q parameter 
(or the proxy, circular velocity to dispersion ratio; $v_{\rm circ}/\sigma$; e.g., \citealt{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}).
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode ($m$\eq0), the dispersion relation for the growth of density perturbation in a rotating turbulent disk
of finite thickness $h$ 
is described by:
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right).
\label{eqn:kappa}
\end{equation}
% \kappa\propto\Omega
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and
the third term describes pressure support.
One can then define the so-called Toomre-Q parameter as:
\begin{equation}
Q\equiv\frac{\sigma\kappa}{\pi G \Sigma}.
\label{eqn:Q}
\end{equation}
For a thin disk, where $kh\ll1$, Equation~\ref{eqn:3Ddisp} reduces to the thin disk approximation, which 
becomes unstable % perturbation grows exponentially in time
for some $k$ when $\omega^2 < 0$ (or equivalently, $Q < 1$; \citealt{Toomre64a}).
Under this condition (i.e., when local gravitational instability 
overcomes support from pressure and rotation), fragmentation {\it can} happen. 
% but further evolution depends on their EOS.
%(e.g., instability impose by angular momentum from gas accretion from the cosmic web and
%satellite galaxies and stellar feedback).
Similarly, for a collisionless fluid, the Toomre Q parameter is defined by:
\begin{equation}
Q\equiv\frac{\sigma\kappa}{3.36 G \Sigma}
\end{equation}
\citep{Toomre64a}.

To account for stabilizing effect of both gas and stars and the non-negligible disk thickness, 
% since thickness of a disk weakens the gravitational forces within the disk plane and lower the values for $Q$
% Romeo:  The effect of thickness is to increase the stability parameter of each 
% component by a factor T, which depends on the ratio of vertical to radial velocity dispersion, following eqn 8 of their paper.
we consider the approximation for the effective two-component $Q_{\rm eff}$ parameter following \citet{Romeo11a} which is 
expressed as follows:
\begin{equation}
Q^{-1}_{\rm eff} = \[ \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{T_{\rm star} Q_{\rm star}} + \frac{1}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \geq T_{\rm gas} Q_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{T_{\rm star} Q_{\rm star}} + \frac{w}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \leq T_{\rm gas} Q_{\rm gas} \\				
				\end{array} 
			    \right,
			    \]
\end{equation}
where $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\rm star} \sigma_{\rm gas}}{\sigma_{\rm star}^2 + \sigma_{\rm gas}^2},
\end{equation}
and $T$ is defined as:
\begin{equation}
T\approx
0.8+0.7\left(\frac{\sigma_{z}}{\sigma_{R}}\right)\,.
\end{equation}

>>>>

\begin{abstract}
We study the dynamical properties of molecular clouds complexes (MC) in a prototypical galaxy (\flower)
and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
We identify MCs using H$_2$ density, 
and compare their masses, sizes, velocity dispersions, gas surface densities, and virial parameters ($\alpha_{\rm vir}$)
with those observed in nearby and \z$\sim$\,2 galaxies.
The MCs span a mass range of $10^{6.5-9}$\,\Msun. Their velocity dispersions and gas surface densities are 
systematically higher than those observed in nearby galaxies but comparable to 
starburst galaxies, with $\sigma\gtrsim$\,100\,\kms.
% Toomre Q
We compute the effective two-component Toomre Q parameter (gas and stars) to examine 
regions that are unstable to local axisymmetric perturbations. 
We further perform virial analysis, which is observable, to examine collapsing regions.
Virial analysis indicates that the MCs in the main disk of \flower are unbound, but its 
substructures have lower virial parameters,
consistent with the notion that collapsing structures result from gravitational instability within globally stable structures.
% , which are supported by turbulence and rotation on large scale.
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, 
which we interpret as result from the weaker stellar feedback
due to the less frequent \SF compared to \flower.
Our results suggest that the more massive and bigger MCs of \flower compared to the Milky Way
manifest from the higher gas mass fraction, surface density, and velocity dispersion.
% , which together sets the scale for fragmentation.
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the
Next Generation Very Large Array (ngVLA)
will furnish scrutinizing data to test our findings and shed light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

In theoretical and observational works, gravitational instability has been widely discussed in the 
context of the Toomre stability criterion \citep{Toomre64a, Goldreich65b}, or the Toomre Q parameter 
(or the proxy, circular velocity to dispersion ratio; $v_{\rm circ}/\sigma$; e.g., \citealt{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}).
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%
% from linear analysis
For {\it axisymmetric} mode ($m$\eq0), the dispersion relation for the growth of density perturbation in a rotating turbulent disk
of finite thickness $h$ 
is described by:
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma^2 k^2,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{R}\frac{d}{dR}\left(R^2\Omega\right).
\label{eqn:kappa}
\end{equation}
% \kappa\propto\Omega
The first term on the RHS of Equation~\ref{eqn:3Ddisp} describes rotation support, whereas the second term describes self-gravity, and
the third term describes pressure support.
One can then define the so-called Toomre-Q parameter as:
\begin{equation}
Q\equiv\frac{\sigma\kappa}{\pi G \Sigma}.
\label{eqn:Q}
\end{equation}
For $kh\ll1$, Equation~\ref{eqn:3Ddisp} reduces to the thin disk approximation, which 
becomes unstable % perturbation grows exponentially in time
for some $k$ when $\omega^2 < 0$ (or equivalently, $Q < 1$; \citealt{Toomre64a}).
Under this condition (i.e., when local gravitational instability 
overcomes support from pressure and rotation), fragmentation {\it can} happen. 
% but further evolution depends on their EOS.
%(e.g., instability impose by angular momentum from gas accretion from the cosmic web and
%satellite galaxies and stellar feedback).
Similarly, for a collisionless fluid, the Toomre Q parameter is defined by:
\begin{equation}
Q\equiv\frac{\sigma\kappa}{3.36 G \Sigma}
\end{equation}
\citep{Toomre64a}.

To account for stabilizing effect of both gas and stars and the non-negligible disk thickness, 
% since thickness of a disk weakens the gravitational forces within the disk plane and lower the values for $Q$
% Romeo:  The effect of thickness is to increase the stability parameter of each 
% component by a factor T, which depends on the ratio of vertical to radial velocity dispersion, following eqn 8 of their paper.
we consider the approximation for the effective two-component $Q_{\rm eff}$ parameter following \citet{Romeo11a, Inoue16a}
which is expressed as follows:
\begin{equation}
Q^{-1}_{\rm eff} = \[ \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{T_{\rm star} Q_{\rm star}} + \frac{1}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \geq T_{\rm gas} Q_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{T_{\rm star} Q_{\rm star}} + \frac{w}{T_{\rm gas} Q_{\rm gas}}}      & & & \mbox{if\ }T_{\rm star} Q_{\rm star} \leq T_{\rm gas} Q_{\rm gas} \\				
				\end{array} 
			    \right,
			    \]
\end{equation}
where $w$ is defined as:
\begin{equation}
w\equiv\frac{2 \sigma_{\rm star} \sigma_{\rm gas}}{\sigma_{\rm star}^2 + \sigma_{\rm gas}^2},
\end{equation}
and $T$ is approximated as:
\begin{equation}
T = \[ \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [0.75em]
			{\dispalystyle 1 + 0.6 \left(\frac{\sigma_z}{\sigam_r}\right)}             & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r 
\\
		\end{array}
		\right.
		\]
\end{equation}
